---
title: "Analisis ANOVAS en Controles para ver tendencias entre especies"
author: "Andrea Chavez"
date: "01-06-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PRELIMINAR ANALYSIS FOR NUMEX RESULTS.
Here are the analysis at species level. 


```{r libraries}
library(dplyr)
library(tidyverse)
library(glmmTMB)
library(DHARMa)
library(ggplot2)
library(scico)
library(lmerTest)
library(lme4)
library(tables)
library(flextable)
library(readxl)
library(modelsummary)
library(knitr)
library(patchwork)
library("kableExtra")
library("openxlsx")
```

```{r Data}
getwd()

data_for_R <- read_excel('data_for_R.xlsx')

data= data_for_R
```

```{r transformando columnas}

### nutrientes para ver cual funciona mejor.
data$logNmass <- log(data$LNC) #
data$logCmass <- log(data$LCC) #
data$logPmass <- log(data$LPC) #


data$SUB = data$Sub #porque abajo esta en mayuscula

#solo dejar las necesarias
data <- dplyr::select(data,
        `name`,
        `genus`,
        `SUB`,#
        `N`,
        `P`,
        `Bloque`,
        `elevation_m`,
    `iWUE (μmol/mol)`,
    `δ18O in ‰ vs DeltaSMOW`,
    `Specific leaf area (cm²/g)`,
    `δ13C (‰ v.s.V-PDB)`,
    `force to punch (kN/m)`, `Leaf area (cm²)`,`dry matter content (mg/g)`, `thickness (mm)`,`N:P`,`Ci/Ca`, Ci,LNC,LPC,LCC,logNmass,logCmass,logPmass,P_area,N_area,C_area) 
#Δ13C_cell
```

# ANOVA
Tendencias generales



```{r ANOVA }
#global ANOVA for control
controles <- data %>%
  filter(SUB == "C")
# Species order
table(controles$name, controles$elevation_m)

controles$elevation_m <- as.factor(controles$elevation_m)
controles$name <- as.factor(controles$name)
  
#anova <- aov( `iWUE (μmol/mol)`~ elevation_m, data = controles)
anova <- aov( `LNC`~ elevation_m, data = controles) #Uses all dataset. Interaction between species and treatment
#Uses all dataset. Interaction between species and treatment
summary(anova)

TukeyHSD(anova, "elevation_m")

shapiro.test(residuals(anova))
```

```{r ANOVA }

library(multcompView)
library(dplyr)
shapiro.test(residuals(anova))
tukey <- TukeyHSD(anova)

# Orden deseado
controles$name <- factor(
  controles$name,
  levels = c(
    "Pouteria torta",
    "Clarisia  racemosa",
    "Myrcia sp nov",
    "Alchornea lojaensis",
    "Weinmania  loxensis",
    "Hedyosmun  purpurascens"
  )
)

#controles <- controles %>%
 # rename(N_P = `N:P`)

controles <- controles %>%
  rename(iWUE = `iWUE (μmol/mol)`)

anova <- aov(iWUE ~ name, data = controles)

tukey <- TukeyHSD(anova)

letters <- multcompLetters4(anova, tukey)

cld <- data.frame(
  name = names(letters$name$Letters),
  letra = letters$name$Letters
)

maximos <- controles %>%
  group_by(name) %>%
  summarise(y = max(iWUE, na.rm = TRUE)) ###edy

cld <- left_join(cld, maximos, by = "name")

ggplot(controles, aes(x = name, y = iWUE)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2) +
  geom_text(
    data = cld,
    aes(x = name, y = y + 0.3, label = letra),  # aumentar este valor
    size = 5,
    fontface = "bold"
  ) +
  theme_bw() +
  labs(
    x = "Species",
    y = "iWUE"
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      face = "italic"
    )
  )
```

Resultados del ANOVA una sola via especie:
### LNC 
Between especies, especies at 1000 m, Clarisia and Pouteria are similar to each other (p-value = 0.20). Weinmania and Hedyosmum although being in the same elevation are different (p-value = 0.0007784). Myrcia and Alchornea are different too. (p=0.008). 
### Narea. At mid elevation the highest Narea. High N levels at low levels are not being traslated into high Narea.Leaves are bigger at low elevations.At high elevations leaves are smaller but it is not traslated into bigger Narea.

### iWUE
Species at 1000 share similar iWUE. Myrcia and Alchornea are slightly different between them. Species at 2000 and 3000 are similar to each other.

### LPC 
Not pattern with elevation.P_area slightly higher at mid elevation with Alchornea taking up more. 
```{r}
library(dplyr)
library(ggplot2)
library(multcompView)
library(patchwork)

# Función para generar cada gráfico
make_plot <- function(data, variable, ylab){

  datos <- data %>%
    rename(valor = all_of(variable))

  anova <- aov(valor ~ name, data = datos)
  tukey <- TukeyHSD(anova)

  letters <- multcompLetters4(anova, tukey)

  cld <- data.frame(
    name = names(letters$name$Letters),
    letra = letters$name$Letters
  )

  maximos <- datos %>%
    group_by(name) %>%
    summarise(y = max(valor, na.rm = TRUE),
              .groups = "drop")

  cld <- left_join(cld, maximos, by = "name")

  ggplot(datos, aes(x = name, y = valor)) +
    geom_boxplot() +
    geom_jitter(width = 0.15, alpha = 0.6, size = 2) +
    geom_text(
      data = cld,
      aes(x = name, y = y + 2.5, label = letra),
      size = 3,
      fontface = "bold"
    ) +
    theme_bw() +
    labs(
      x = NULL,
      y = ylab
    ) +
    theme(
      axis.text.x = element_text(
        angle = 45,
        hjust = 1,
        face = "italic"
      )
    )
}

# Crear gráficos
p1 <- make_plot(controles, "iWUE", "iWUE (µmol mol⁻¹)")
p2 <- make_plot(controles, "N_P", "N:P")
p3 <- make_plot(controles, "LNC", "Leaf N (mg/g)")
p4 <- make_plot(controles, "LPC", "Leaf P (mg/g)")

p1 <- p1 +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

p2 <- p2 +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )


((p1 + p2) /
 (p3 + p4)) +plot_annotation(tag_levels = "A")
```


```{r}

library(tidyr)
library(ggplot2)

datos_long <- controles %>%
  select(name, iWUE, N_P, LNC, LPC) %>%
  pivot_longer(
    cols = c(iWUE, N_P, LNC, LPC),
    names_to = "Variable",
    values_to = "Valor"
  )

ggplot(datos_long, aes(name, Valor)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.6) +
  facet_wrap(~ Variable, ncol = 2, scales = "free_y") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      face = "italic"
    )
  ) +
  labs(x = "Species", y = NULL)
```

