---
title: "Analisis ANOVAS en Controles para ver tendencias entre especies"
author: "Andrea Chavez"
date: "01-06-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PRELIMINAR ANALYSIS FOR NUMEX RESULTS.
Here are the analysis at species level. 


```{r libraries}
library(dplyr)
library(tidyverse)
library(glmmTMB)
library(DHARMa)
library(ggplot2)
library(scico)
library(lmerTest)
library(lme4)
library(tables)
library(flextable)
library(readxl)
library(modelsummary)
library(knitr)
library(patchwork)
library("kableExtra")
library("openxlsx")
```

```{r Data}
getwd()

data_for_R <- read_excel('data_for_R.xlsx')

data= data_for_R
```

```{r transformando columnas}

### nutrientes para ver cual funciona mejor.
data$logNmass <- log(data$LNC) #
data$logCmass <- log(data$LCC) #
data$logPmass <- log(data$LPC) #


data$SUB = data$Sub #porque abajo esta en mayuscula

#solo dejar las necesarias
data <- dplyr::select(data,
        `name`,
        `genus`,
        `SUB`,#
        `N`,
        `P`,
        `Bloque`,
        `elevation_m`,
    `iWUE (μmol/mol)`,
    `δ18O in ‰ vs DeltaSMOW`,
    `Specific leaf area (cm²/g)`,
    `δ13C (‰ v.s.V-PDB)`,
    `force to punch (kN/m)`, `Leaf area (cm²)`,`dry matter content (mg/g)`, `thickness (mm)`,`N:P`,`Ci/Ca`, Ci,LNC,LPC,LCC,logNmass,logCmass,logPmass,P_area,N_area,C_area) 
#Δ13C_cell
```

# ANOVA
Tendencias generales

```{r ANOVA }
#global ANOVA for control
controles <- data %>%
  filter(SUB == "C")
# Species order
table(controles$name, controles$elevation_m)

controles$elevation_m <- as.factor(controles$elevation_m)
controles$name <- as.factor(controles$name)
  
#anova <- aov( `iWUE (μmol/mol)`~ elevation_m, data = controles)
anova_LNC <- aov( `LNC`~ name, data = controles) #Uses all dataset.
anova_LPC <- aov( `LPC`~ name, data = controles) #Uses all dataset.
anova_iWUE <- aov(`iWUE (μmol/mol)` ~ name, data = controles)
anova_Pa <- aov( `P_area`~ name, data = controles) #Uses all dataset.
anova_Na <- aov( `N_area`~ name, data = controles) #Uses all dataset.
anova_d13C <- aov(`δ13C (‰ v.s.V-PDB)` ~ name, data = controles)

#Uses all dataset. Interaction between species and treatment
summary(anova_LNC)
summary(anova_LPC)
summary(anova_iWUE)
summary(anova_Pa)
summary(anova_Na)
summary(anova_d13C)

TukeyHSD(anova_LNC, "name")
TukeyHSD(anova_LPC, "name")
TukeyHSD(anova_iWUE, "name")
TukeyHSD(anova_Pa, "name")
TukeyHSD(anova_Na, "name")
TukeyHSD(anova_d13C, "name")

shapiro.test(residuals(anova_LNC))
shapiro.test(residuals(anova_LPC))
shapiro.test(residuals(anova_iWUE))
shapiro.test(residuals(anova_Pa))
shapiro.test(residuals(anova_Na))
shapiro.test(residuals(anova_d13C))
```
# Grafico individual
```{r grafico individual }

library(multcompView)
library(dplyr)

# Orden deseado
controles$name <- factor(
  controles$name,
  levels = c(
    "Pouteria torta",
    "Clarisia  racemosa",
    "Myrcia sp nov",
    "Alchornea lojaensis",
    "Weinmania  loxensis",
    "Hedyosmun  purpurascens"
  )
)

#controles <- controles %>%
 # rename(N_P = `N:P`)

controles <- controles %>%
  rename(iWUE = `iWUE (μmol/mol)`)
controles <- controles %>%
  rename(d13C = `δ13C (‰ v.s.V-PDB)`)

letters <- multcompLetters4(anova, tukey)

cld <- data.frame(
  name = names(letters$name$Letters),
  letra = letters$name$Letters
)

maximos <- controles %>%
  group_by(name) %>%
  summarise(y = max(iWUE, na.rm = TRUE)) ###edy

cld <- left_join(cld, maximos, by = "name")

ggplot(controles, aes(x = name, y = iWUE)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2) +
  geom_text(
    data = cld,
    aes(x = name, y = y + 0.3, label = letra),  # aumentar este valor
    size = 5,
    fontface = "bold"
  ) +
  theme_bw() +
  labs(
    x = "Species",
    y = "iWUE"
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      face = "italic"
    )
  )
```

Resultados del ANOVA una sola via especie:
LNC 
Between especies, especies at 1000 m, Clarisia and Pouteria are similar to each other (p-value = 0.20). Weinmania and Hedyosmum although being in the same elevation are different (p-value = 0.0007784). Myrcia and Alchornea are different too. (p=0.008). 
### Narea. At mid elevation the highest Narea. High N levels at low levels are not being traslated into high Narea.Leaves are bigger at low elevations.At high elevations leaves are smaller but it is not traslated into bigger Narea.

iWUE
Species at 1000 share similar iWUE. Myrcia and Alchornea are slightly different between them. Species at 2000 and 3000 are similar to each other.

LPC 
Not pattern with elevation.P_area slightly higher at mid elevation with Alchornea taking up more. 
# graficos
```{r}
library(dplyr)
library(ggplot2)
library(multcompView)
library(patchwork)


make_plot <- function(data, variable, ylab, ylim = NULL, label_offset = NULL){
  datos <- data %>%
    rename(valor = all_of(variable))
  anova <- aov(valor ~ name, data = datos)
  tukey <- TukeyHSD(anova)
  letters <- multcompLetters4(anova, tukey)
  cld <- data.frame(
    name = names(letters$name$Letters),
    letra = letters$name$Letters
  )
  
  # si no se especifica offset, usar 2.5% del rango de los datos
  if (is.null(label_offset)) {
    rango <- max(datos$valor, na.rm = TRUE) - min(datos$valor, na.rm = TRUE)
    label_offset <- rango * 0.05
  }
  
  maximos <- datos %>%
    group_by(name) %>%
    summarise(y = max(valor, na.rm = TRUE), .groups = "drop")
  cld <- left_join(cld, maximos, by = "name")
  
  p <- ggplot(datos, aes(x = name, y = valor)) +
    geom_boxplot() +
    geom_jitter(width = 0.15, alpha = 0.6, size = 2) +
    geom_text(
      data = cld,
      aes(x = name, y = y + label_offset, label = letra),
      size = 3,
      fontface = "bold"
    ) +
    theme_bw(base_size = 12) +
    labs(x = NULL, y = ylab) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, face = "italic")
    )
  
  if (!is.null(ylim)) {
    p <- p + coord_cartesian(ylim = ylim)
  }
  
  return(p)
}

# Crear gráficos
p1 <- make_plot(controles, "iWUE", "iWUE (µmol mol⁻¹)")
p2 <- make_plot(controles, "d13C", "δ13C (‰ v.s.V-PDB)")
p3 <- make_plot(controles, "LNC", "Leaf N (mg/g)")
p4 <- make_plot(controles, "N_area", "N area")

p5 <- make_plot(controles, "LPC", "Leaf P (mg/g)")
p6 <- make_plot(controles, "P_area", "P_area", ylim = c(0, 0.5), label_offset = 0.03)

p7 <- make_plot(controles, "N:P", "N:P")

p1 <- p1 +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

p2 <- p2 +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

p3 <- p3 +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

p4 <- p4 +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )


figura_final <- ((p1 + p2) /
                  (p3 + p4) /
                  (p5 + p6)) +
  plot_annotation(tag_levels = "A")

ggsave(
  filename = "figura_controles.png",
  plot = figura_final,
  width = 180,     # mm (A4 = 210mm, dejamos márgenes)
  height = 240,    # mm (A4 = 297mm)
  units = "mm",
  dpi = 600,       # 300 mínimo, 600 recomendado para revistas
  bg = "white"
)
```


```{r}

library(tidyr)
library(ggplot2)

datos_long <- controles %>%
  select(name, iWUE, "N:P", LNC, LPC) %>%
  pivot_longer(
    cols = c(iWUE,"N:P", LNC, LPC),
    names_to = "Variable",
    values_to = "Valor"
  )

ggplot(datos_long, aes(name, Valor)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.6) +
  facet_wrap(~ Variable, ncol = 2, scales = "free_y") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      face = "italic"
    )
  ) +
  labs(x = "Species", y = NULL)
```

