---
title: "Analisis per specie"
author: "Andrea Chavez"
date: "15/12/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PRELIMINAR ANALYSIS FOR NUMEX RESULTS.
Here are the analysis at species level. 


```{r libraries}


library(tidyverse)
library(glmmTMB)
library(DHARMa)
library(ggplot2)
library(scico)
library(lmerTest)
library(lme4)
library(tables)
library(dplyr)
library(flextable)
library(readxl)
library(modelsummary)
library(knitr)
library(patchwork)

```

```{r Data}
getwd()

data_for_R <- read_excel('data_for_R.xlsx')

data = data_for_R
```

```{r transformando columnas}

# Crear columnas binarias manualmente para añadir fertilizacion
#columnas binarias

data$logNmass <- log(data$`[N]`)
data$logNarea <- log(data$Narea)

data <- data %>%
  mutate(d15N_centered = scale(`δ15N (‰ v.s.AIR)`, center = TRUE, scale = FALSE)[,1])

#adding the blocks ONLY WORKS FOR CANU Y BONU
#data <- data %>%
#  mutate(
#    Bloque = case_when(
#      Plot %in% 1:4   ~ 1,
##      Plot %in% 5:8   ~ 2,
#      Plot %in% 9:12  ~ 3,
##      Plot %in% 13:16 ~ 4,
#      TRUE ~ NA_real_
#    )
#  )
```



```{r function for species plot}
# Order Sub
#data$Sub <- factor(data$Sub, levels = c("C", "N", "P", "NP"))

# Species order
species_order <- c(
  "Pouteria torta",
  "Alchornea lojaensis",
  "Weinmania  loxensis",
  "Clarisia  racemosa",
  "Myrcia sp nov",
  "Hedyosmun  purpurascens"
)

data$name <- factor(data$name, levels = species_order)

# Elevation map
elevation_map <- c(
  "Pouteria torta" = "1000 m",
  "Alchornea lojaensis" = "2000 m",
  "Weinmania  loxensis" = "3000 m",
  "Clarisia  racemosa" = "1000 m",
  "Myrcia sp nov" = "2000 m",
  "Hedyosmun  purpurascens" = "3000 m"
)

# ===============================
#   MAIN FUNCTION
# ===============================
make_species_plots <- function(yvar, y_label = yvar, y_limits = c(0, 150)) {
  
  # dynamic column reference
  ycol <- rlang::sym(yvar)
  
  make_one_plot <- function(species_name) {
    
    df <- data[data$name == species_name, ]
  
    # mean of control
    mean_control <- mean(df[df$Sub == "C", ][[yvar]], na.rm = TRUE)
    
    ggplot(df, aes(x = Sub, y = !!ycol, colour = Sub)) +
      
      geom_boxplot(aes(group = Sub),
                   colour = "black",
                   fill = "grey90",
                   alpha = 0.5) +
      
      geom_hline(yintercept = mean_control,
                 colour = "black",
                 linetype = "dotted",
                 linewidth = 1) +
      
      geom_jitter(width = 0.2, alpha = 0.6) +
      
      annotate("label",
               x = 4, y = y_limits[2] - 1,
               label = elevation_map[[species_name]],
               fill = "white",
               colour = "black",
               size = 3) +
      
      scale_y_continuous(limits = y_limits) +
      
      labs(title = bquote(italic(.(species_name))),
           x = "",
           y = y_label) +
      
      theme_bw() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
      }
  
  # Build 6 plots
  p_list <- lapply(species_order, make_one_plot)
  
  final_plot <- ((p_list[[1]] | p_list[[2]] | p_list[[3]]) /
                   (p_list[[4]] | p_list[[5]] | p_list[[6]])) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
  
#ggsave(filename= paste0("species_boxplots_", y_label,".png"),
#       plot = final_plot,
#       width = 10, height = 10, units = "in", dpi = 300)

  return(final_plot)
}
```


```{r plot}

plot_iwue <- make_species_plots("iWUE (μmol/mol)",
                                 y_label = "iWUE (μmol/mol)",
                               y_limits = c(25,120))

plot_18O <- make_species_plots("δ18O in ‰ vs DeltaSMOW",
                                y_label = "δ18O in ‰ vs DeltaSMOW",
                                y_limits = c(25,35))
plot_15N <- make_species_plots("δ15N (‰ v.s.AIR)",
                                y_label = "δ15N",
                                y_limits = c(-8,6))
plot_CN <- make_species_plots("C:N",
                                y_label = "CN",
                                y_limits = c(15,60))
plot_N <- make_species_plots("[N]",
                                y_label = "[N]",
                                y_limits = c(0.005,0.025))

plot_Narea <- make_species_plots("Narea",
                                y_label = "Narea",
                                y_limits = c(0,0.5))

```
The graphic for C13/o18 are created in python.
```{r 13C/18O}
library(ggplot2)

ggplot(data, aes(x = `δ18O in ‰ vs DeltaSMOW`, 
                 y = `δ13C (‰ v.s.V-PDB)`,colour = Sub)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ name, ncol = 3) +
  theme_bw() +
  labs(
    x = expression(delta^{18}*O),
    y = expression(delta^{13}*C),
    title = "Relationship between δ13C and δ18O per species"
  )


```


```{r histograms}

species_vector <- c(
  "Pouteria torta",
  "Alchornea lojaensis",
  "Weinmania  loxensis",
  "Clarisia  racemosa",
  "Myrcia sp nov",
  "Hedyosmun  purpurascens"
) 
plot_6_species_histograms <- function(data, species_vector, var_name = "iWUE") {
  
  # set grid: 2 rows × 3 columns
  par(mfrow = c(2, 3))
  
  for (sp in species_vector) {
    df <- subset(data, name == sp)
    
    # check if variable exists
    
    x <- df[[var_name]]
    
    # if no data
    if (length(x) == 0 || !is.numeric(x)) {
      plot.new()
      title(main = paste("No data for:", sp))
      next
    }
    
    hist(x,
         main = paste("Histogram:", sp),
         xlab = var_name)
  }
}
plot_6_species_histograms(data, species_vector, var_name = "iWUE (μmol/mol)") #add the different variables here.


```

# ANOVA

```{r ANOVA }
#global ANOVA for all the samples
#variables = δ18O in ‰ vs DeltaSMOW`
  
anova_species_treat <- aov( `iWUE (μmol/mol)`~ name * Sub, data = data) #Uses all dataset. Interaction between species and treatment
summary(anova_species_treat)
#Each specie is giving different responses.
```

### ANOVA 1000

```{r 1000 m}

### Bombuscaro ####
# Pouteria for isotopes results ##
# other suggested model 
datos_Pouteria <- data %>%
filter(data$genus == "Pouteria")

datos_Pouteria_filtered <- datos_Pouteria[datos_Pouteria$`iWUE (μmol/mol)` < 100, ]

anova_model_iWUE <- aov(`iWUE (μmol/mol)` ~ N * P, data = datos_Pouteria_filtered)#keep all the rest
anova_model_18O <- aov(`δ18O in ‰ vs DeltaSMOW` ~ N * P, data = datos_Pouteria)
anova_model_15N <- aov(`δ15N (‰ v.s.AIR)` ~ N * P, data = datos_Pouteria)
anova_model_CN <- aov(`Narea` ~ N * P, data = datos_Pouteria) 

#Diagnostico del modelo
#par(mfrow = c(2, 2))
#plot(anova_model_iWUE)
#par(mfrow = c(1, 1))

#### Bombuscaro ####
### Clarisia ####
datos_Clarisia <- data %>%
filter(data$genus == "Clarisia")

anova_model_15N_cr <- aov(`δ15N (‰ v.s.AIR)` ~ N * P, data = datos_Clarisia)

anova_model_18O_cr <- aov(`δ18O in ‰ vs DeltaSMOW` ~ N * P, data = datos_Clarisia)

anova_model_iWUE_cr <- aov(`iWUE (μmol/mol)` ~ N * P, data = datos_Clarisia)
anova_model_CN_cr <- aov(`Narea` ~ N * P, data = datos_Clarisia)

## Para la Clarisia,, no hay ninguna diferencia en iWUE. En 18O, un poco efecto de N pero hubieron muy pocas muestras para adelantar conclusiones. En N, incrementa 18O. Transpira mas.In this case N has an effect on 15N for Clarisia.Lower 15N values in the N group.


```
### ANOVA 2000
```{r anova 2000m}

##### San Francisco #######
## Alchornea ###
datos_al <- data %>%
filter(data$name == "Alchornea lojaensis")

datos_al<- datos_al%>% #the blocks change because there is a difference between.
  mutate(
    Bloque = case_when(
      Plot %in% 1:4   ~ 1,
      Plot %in% 6:9   ~ 2,
      Plot %in% 11:14  ~ 3,
      Plot %in% 16:19 ~ 4,
      TRUE ~ NA_real_
    )
  )

datos_al <- datos_al[!is.nan(datos_al$`δ18O in ‰ vs DeltaSMOW`), ]


anova_model_15N_al <- aov(`δ15N (‰ v.s.AIR)` ~ N * P, data = datos_al)
anova_model_18O_al<- aov(`δ18O in ‰ vs DeltaSMOW` ~ N * P, data = datos_al)
anova_model_iWUE_al <- aov(`iWUE (μmol/mol)` ~ N * P, data = datos_al)
anova_model_CN_al <- aov(`Narea` ~ N * P, data = datos_al)

datos_myrcia <- data %>%
filter(data$name == "Myrcia sp nov")

#datos_myrcia<- datos_myrcia %>%
#  mutate(
#    Bloque = case_when(
#      Plot %in% 1:4   ~ 1,
#      Plot %in% 6:9   ~ 2,
#      Plot %in% 11:14  ~ 3,
#      Plot %in% 16:19 ~ 4,
#      TRUE ~ NA_real_
#    )
#  )

datos_myrcia <- datos_myrcia[!is.nan(datos_myrcia$`δ18O in ‰ vs DeltaSMOW`), ]


anova_model_15N_m <- aov(`δ15N (‰ v.s.AIR)` ~ N * P, data = datos_myrcia)
anova_model_18O_m <- aov(`δ18O in ‰ vs DeltaSMOW` ~ N * P, data = datos_myrcia)
anova_model_iWUE_m <- aov(`iWUE (μmol/mol)` ~ N * P, data = datos_myrcia)
anova_model_CN_m <- aov(`Narea` ~ N * P, data = datos_myrcia)

```
### ANOVA 3000


```{r anova 3000 m}
#### Cajanuma #####
### Weinmania ####

## For Weinmania. No significant effect on iWUE, marginal effect of P. Lower iWUE in P than C. Then for 18O and 15N no difference between treatments. 

datos_Weinmania <- data %>%
filter(data$genus == "Weinmania")


anova_model_15N_Wl<- aov(`δ15N (‰ v.s.AIR)` ~ N * P, data = datos_Weinmania)

anova_model_18O_Wl <- aov(`δ18O in ‰ vs DeltaSMOW` ~ N * P, data = datos_Weinmania)

anova_model_iWUE_Wl <- aov(`iWUE (μmol/mol)` ~ N * P, data = datos_Weinmania)
anova_model_CN_Wl <- aov(`Narea` ~ N * P, data = datos_Weinmania)

### Hedyosmum purpurascens ####

datos_Hp <- data %>%
filter(data$genus == "Hedyosmun")

anova_model_15N_Hp <- aov(`δ15N (‰ v.s.AIR)` ~ N * P, data = datos_Hp)

anova_model_18O_Hp <- aov(`δ18O in ‰ vs DeltaSMOW` ~ N * P, data = datos_Hp)

anova_model_iWUE_Hp <- aov(`iWUE (μmol/mol)` ~ N * P, data = datos_Hp)
anova_model_CN_Hp <- aov(`Narea` ~ N * P, data = datos_Hp)

#Hedyosmum, no significant effect of any nutrient on these ecophysiological parameters. 
```
#### TABLA ANOVA 1000
```{r TABLAS ANOVA 1000 }

mods_1000 <- list(
  "Pouteria iWUE" = anova_model_iWUE,
  "δ18O" = anova_model_18O,
  "δ15N" = anova_model_15N,
  "C:N"  = anova_model_CN,
  "Clarisia iWUE " = anova_model_iWUE_cr,
  "δ18O" = anova_model_18O_cr,
  "δ15N" = anova_model_15N_cr,
  "C:N"  = anova_model_CN_cr
)

modelsummary(
  mods_1000,
  statistic = "({statistic}, p = {p.value})",
  stars = c('*' = .05, '**' = .01, '***' = .001),
  fmt = 2,
  output = "markdown",
  title = "ANOVA results for *1000 species*")
```
Pouteria: 
iWUE relevant NxP
d18O N, P, no the interaction
d15N .
C:N only with P.

Clarisia:
d18I for N,
d15N for N
C:N for N
Clarisia N is important.

#### TABLA ANOVA 2000
```{r TABLAS ANOVA 2000}

mods_2000 <- list(
  "Alchornea iWUE" = anova_model_iWUE_al,
  "δ18O" = anova_model_18O_al,
  "δ15N" = anova_model_15N_al,
  "C:N"  = anova_model_CN_al,
  "Myrcia iWUE " = anova_model_iWUE_m,
  "δ18O" = anova_model_18O_m,
  "δ15N" = anova_model_15N_m,
  "C:N"  = anova_model_CN_m
)
modelsummary(
  mods_2000,
  statistic = "({statistic}, p = {p.value})",
  stars = c('*' = .05, '**' = .01, '***' = .001),
  fmt = 2,
  output = "markdown",
  title = "ANOVA results for 2000 species*")

```
Alchornea or Myrcia. ## HELP error with a database
C:N influenced by N (p=0.018)

#### TABLA ANOVA 3000

```{r TABLAS ANOVA 3000}

mods_3000 <- list(
  "Weinmannia – iWUE" = anova_model_iWUE_Wl,
  "Weinmannia – δ18O" = anova_model_18O_Wl,
  "Weinmannia – δ15N" = anova_model_15N_Wl,
  "Weinmannia – C:N"  = anova_model_CN_Wl,
  "Hedyosmum – iWUE"  = anova_model_iWUE_Hp,
  "Hedyosmum – δ18O"  = anova_model_18O_Hp,
  "Hedyosmum – δ15N"  = anova_model_15N_Hp,
  "Hedyosmum – C:N"   = anova_model_CN_Hp
)


modelsummary(
  mods_3000,
  statistic = "({statistic}, p = {p.value})",
  stars = c('*' = .05, '**' = .01, '***' = .001),
  fmt = 2,
  output = "markdown",
  title = "ANOVA results for *3000 species*")
  
```
# Modelos lineares mixtos 
## MODELOS SIN INTERACCION 
Se elimina de todos N+P por redundancia

```{r IGNORAR modelos lineares por especie con N+P}

### Bombuscaro ####
# Pouteria for isotopes results ##
# other suggested model 

# Reorder factors so "Control" is the reference

# Fit model
mod_iWUE_pt <- lmer(`iWUE (μmol/mol)` ~ N+P+ (1 | Bloque/Plot), data = datos_Pouteria)
mod_O18_pt <- lmer(`δ18O in ‰ vs DeltaSMOW` ~ N+P+ (1 | Bloque/Plot), data = datos_Pouteria)
mod_N15_pt <- lmer(`δ15N (‰ v.s.AIR)` ~ N+P+ (1 | Bloque/Plot), data = datos_Pouteria)
mod_CN_pt <- lmer(`C:N` ~ N+P+ (1 | Bloque/Plot), data = datos_Pouteria)

mod_iWUE_cr <- lmer(`iWUE (μmol/mol)` ~ N+P + (1 | Bloque/Plot), data = datos_Clarisia)
mod_O18_cr<- lmer(`δ18O in ‰ vs DeltaSMOW` ~N+P+ (1 | Bloque/Plot), data = datos_Clarisia)
mod_N15_cr<- lmer(`δ15N (‰ v.s.AIR)` ~ N+P+(1 | Bloque/Plot), data = datos_Clarisia)
mod_CN_cr<- lmer(`C:N` ~N+P+  (1 | Bloque/Plot), data = datos_Clarisia)

mod_iWUE_al <- lmer(`iWUE (μmol/mol)` ~ N+P + (1 | Bloque/Plot), data = datos_al)
mod_O18_al<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ N+P+ (1 | Bloque/Plot), data = datos_al)
mod_N15_al<- lmer(`δ15N (‰ v.s.AIR)` ~ N+P+ (1 | Bloque/Plot), data = datos_al)
mod_CN_al<- lmer(`C:N` ~N+P+ (1 | Bloque/Plot), data = datos_al)

mod_iWUE_my <- lmer(`iWUE (μmol/mol)` ~ N+P+(1 | Bloque/Plot), data = datos_myrcia)
mod_O18_my<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ N+P+(1 | Bloque/Plot), data = datos_myrcia)
mod_N15_my<- lmer(`δ15N (‰ v.s.AIR)` ~ N+P+ (1 | Bloque/Plot), data = datos_myrcia)
mod_CN_my<- lmer(`C:N` ~ N+P+(1 | Bloque/Plot), data = datos_myrcia)


mod_iWUE_hp <- lmer(`iWUE (μmol/mol)` ~N+P + (1 | Bloque/Plot), data = datos_Hp)
mod_O18_hp<- lmer(`δ18O in ‰ vs DeltaSMOW` ~N+P+ (1 | Bloque/Plot), data = datos_Hp)
mod_N15_hp<- lmer(`δ15N (‰ v.s.AIR)` ~ N+P+(1 | Bloque/Plot), data = datos_Hp)
mod_CN_hp<- lmer(`C:N` ~ N+P+(1 | Bloque/Plot), data = datos_Hp)


mod_iWUE_wl <- lmer(`iWUE (μmol/mol)` ~ N+P+ (1 | Bloque/Plot), data = datos_Weinmania)
mod_O18_wl<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ N+P+ (1 | Bloque/Plot), data = datos_Weinmania)
mod_N15_wl<- lmer(`δ15N (‰ v.s.AIR)` ~ N+P+ (1 | Bloque/Plot), data = datos_Weinmania)
mod_CN_wl<- lmer(`C:N` ~ N + P + (1 | Bloque/Plot), data = datos_Weinmania)

#mod_iWUE_2 <- lmer(`iWUE (μmol/mol)` ~ Sub + (1 | Plot), data = datos_Pouteria) #HELP: Why

models  <- list(
  "Pouteria iWUE" = mod_iWUE_pt,
  "Pouteria O18"  = mod_O18_pt,
  "Pouteria N15"  = mod_N15_pt,
  "Pouteria CN"   = mod_CN_pt,

  "Clarisia iWUE" = mod_iWUE_cr,
  "Clarisia O18"  = mod_O18_cr,
  "Clarisia N15"  = mod_N15_cr,
  "Clarisia CN"   = mod_CN_cr,

  "Alchornea iWUE" = mod_iWUE_al,
  "Alchornea O18"  = mod_O18_al,
  "Alchornea N15"  = mod_N15_al,
  "Alchornea CN"   = mod_CN_al,

  "Myrcia iWUE" = mod_iWUE_my,
  "Myrcia O18"  = mod_O18_my,
  "Myrcia N15"  = mod_N15_my,
  "Myrcia CN"   = mod_CN_my,

  "Hedyosmum iWUE" = mod_iWUE_hp,
  "Hedyosmum O18"  = mod_O18_hp,
  "Hedyosmum N15"  = mod_N15_hp,
  "Hedyosmum CN"   = mod_CN_hp,

  "Weinmania iWUE" = mod_iWUE_wl,
  "Weinmania O18"  = mod_O18_wl,
  "Weinmania N15"  = mod_N15_wl,
  "Weinmania CN"   = mod_CN_wl)
# exportar tabla
tab <- modelsummary(
  models,
  output = "data.frame",
  stars = TRUE,
  gof_map = c("nobs", "r2.marginal", "aic")
)

# transponer
tab_t <- as.data.frame(t(tab))

# fijar nombres de columna usando la primera fila
colnames(tab_t) <- tab_t[1, ]
tab_t <- tab_t[-1, ]  # quitar fila original de nombres

# mostrar tabla
kable(tab_t)

```
The interaction is only relevant for one species. The Pouteris and Myrcia N15.

## MODELOS CON DUMMY BLOQUE E INTERACCION
```{r modelos DUMMY Y BLOQUE}

mod_iWUE_pt <- lmer(`iWUE (μmol/mol)` ~ N*P + (1 | Bloque), data = datos_Pouteria)
mod_O18_pt <- lmer(`δ18O in ‰ vs DeltaSMOW` ~ N*P + (1 | Bloque), data = datos_Pouteria)
mod_N15_pt <- lmer(`δ15N (‰ v.s.AIR)` ~ N*P + (1 | Bloque), data = datos_Pouteria)
mod_CN_pt <- lmer(`C:N` ~ N*P+ (1 | Bloque), data = datos_Pouteria)
mod_N_pt <- lmer(`[N]` ~ N*P+ (1 | Bloque), data = datos_Pouteria)
mod_Narea_pt <- lmer(`Narea` ~ N*P+ (1 | Bloque), data = datos_Pouteria)
mod_logNarea_pt <- lmer(`logNarea` ~ N*P+ (1 | Bloque), data = datos_Pouteria)
mod_logNmass_pt <- lmer(`logNmass` ~ N*P+ (1 | Bloque), data = datos_Pouteria)

mod_iWUE_cr <- lmer(`iWUE (μmol/mol)` ~ N*P + (1 | Bloque), data = datos_Clarisia)
mod_O18_cr<- lmer(`δ18O in ‰ vs DeltaSMOW` ~N*P  + (1 | Bloque), data = datos_Clarisia)
mod_N15_cr<- lmer(`δ15N (‰ v.s.AIR)` ~ N*P + (1 | Bloque), data = datos_Clarisia)
mod_CN_cr<- lmer(`C:N` ~N*P+ (1 | Bloque), data = datos_Clarisia)
mod_N_cr <- lmer(`[N]` ~ N*P+ (1 | Bloque), data = datos_Clarisia)
mod_Narea_cr <- lmer(`Narea` ~ N*P+ (1 | Bloque), data = datos_Clarisia)
mod_logNarea_cr <- lmer(`logNarea` ~ N*P+ (1 | Bloque), data = datos_Clarisia)
mod_logNmass_cr <- lmer(`logNmass` ~ N*P+ (1 | Bloque), data = datos_Clarisia)

mod_iWUE_al <- lmer(`iWUE (μmol/mol)` ~ N*P + (1 | Bloque), data = datos_al)
mod_O18_al<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ N*P + (1 | Bloque), data = datos_al)
mod_N15_al<- lmer(`δ15N (‰ v.s.AIR)` ~ N*P+  (1 | Bloque), data = datos_al)
mod_CN_al<- lmer(`C:N` ~N*P + (1 | Bloque), data = datos_al)
mod_N_al <- lmer(`[N]` ~ N*P+ (1 | Bloque), data = datos_al)
mod_Narea_al <- lmer(`Narea` ~ N*P+ (1 | Bloque), data = datos_al)
mod_logNarea_al <- lmer(`logNarea` ~ N*P+ (1 | Bloque), data = datos_al)
mod_logNmass_al <- lmer(`logNmass` ~ N*P+ (1 | Bloque), data = datos_al)

mod_iWUE_my <- lmer(`iWUE (μmol/mol)` ~ N*P + (1 | Bloque), data = datos_myrcia)
mod_O18_my<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ N*P+ (1 | Bloque), data = datos_myrcia)
mod_N15_my<- lmer(`δ15N (‰ v.s.AIR)` ~N*P+ (1 | Bloque), data = datos_myrcia)
mod_CN_my<- lmer(`C:N` ~ N*P+ (1 | Bloque), data = datos_myrcia)
mod_N_my <- lmer(`[N]` ~ N*P+ (1 | Bloque), data = datos_myrcia)
mod_Narea_my <- lmer(`Narea` ~ N*P+ (1 | Bloque), data = datos_myrcia)
mod_logNarea_my <- lmer(`logNarea` ~ N*P+ (1 | Bloque), data = datos_myrcia)
mod_logNmass_my <- lmer(`logNmass` ~ N*P+ (1 | Bloque), data = datos_myrcia)

mod_iWUE_hp <- lmer(`iWUE (μmol/mol)` ~N*P+ (1 | Bloque), data = datos_Hp)
mod_O18_hp<- lmer(`δ18O in ‰ vs DeltaSMOW` ~N*P+  (1 | Bloque), data = datos_Hp)
mod_N15_hp<- lmer(`δ15N (‰ v.s.AIR)` ~ N*P+(1 | Bloque), data = datos_Hp)
mod_CN_hp<- lmer(`C:N` ~ N*P + (1 | Bloque), data = datos_Hp)
mod_N_hp <- lmer(`[N]` ~ N*P+ (1 | Bloque), data = datos_Hp)
mod_Narea_hp <- lmer(`Narea` ~ N*P+ (1 | Bloque), data = datos_Hp)
mod_logNarea_hp <- lmer(`logNarea` ~ N*P+ (1 | Bloque), data = datos_Hp)
mod_logNmass_hp <- lmer(`logNmass` ~ N*P+ (1 | Bloque), data = datos_Hp)

mod_iWUE_wl <- lmer(`iWUE (μmol/mol)` ~ N*P + (1 | Bloque), data = datos_Weinmania)
mod_O18_wl<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ N*P + (1 | Bloque), data = datos_Weinmania)
mod_N15_wl<- lmer(`δ15N (‰ v.s.AIR)` ~ N*P + (1 | Bloque), data = datos_Weinmania)
mod_CN_wl<- lmer(`C:N` ~ N + P + (1 | Bloque), data = datos_Weinmania)
mod_N_wl <- lmer(`[N]` ~ N*P+ (1 | Bloque), data = datos_Weinmania)
mod_Narea_wl <- lmer(`Narea` ~ N*P+ (1 | Bloque), data = datos_Weinmania)
mod_logNarea_wl <- lmer(`logNarea` ~ N*P+ (1 | Bloque), data = datos_Weinmania)
mod_logNmass_wl <- lmer(`logNmass` ~ N*P+ (1 | Bloque), data = datos_Weinmania)


## Creates a list with all the models
models  <- list(
  "Pouteria iWUE" = mod_iWUE_pt,
  "Pouteria O18"  = mod_O18_pt,
  #"Pouteria N15"  = mod_N15_pt,
  #"Pouteria CN"   = mod_CN_pt,
  "Pouteria Nmass"=mod_N_pt,
  "Pouteria Narea"=mod_Narea_pt,
  "Pouteria logNmass"= mod_logNmass_pt,
  "Pouteria logNarea"= mod_logNarea_wl,

  "Clarisia iWUE" = mod_iWUE_cr,
  "Clarisia O18"  = mod_O18_cr,
  #"Clarisia N15"  = mod_N15_cr,
  #"Clarisia CN"   = mod_CN_cr,
  "Clarisia Nmass" = mod_N_cr,
  "Clarisia Narea" = mod_Narea_cr,
  "Clar logNmass"= mod_logNmass_cr,
  "Clar logNarea"= mod_logNarea_cr,


  "Alchornea iWUE" = mod_iWUE_al,
  "Alchornea O18"  = mod_O18_al,
  #"Alchornea N15"  = mod_N15_al,
  #"Alchornea CN"   = mod_CN_al,
  "Alchornea Nmass"=mod_N_al,
  "Alchornea Narea"=mod_Narea_al,
    "Alchornea logNmass"= mod_logNmass_al,
  "Alchornea logNarea"= mod_logNarea_al,


  "Myrcia iWUE" = mod_iWUE_my,
  "Myrcia O18"  = mod_O18_my,
  #"Myrcia N15"  = mod_N15_my,
  #"Myrcia CN"   = mod_CN_my,
  "Myrcia Nmass" = mod_N_my,
  "Myrcia Narea"  = mod_Narea_my,
    "Myrcia logNmass"= mod_logNmass_my,
  "Myrcia logNarea"= mod_logNarea_my,


  "Hedyosmum iWUE" = mod_iWUE_hp,
  "Hedyosmum O18"  = mod_O18_hp,
  #"Hedyosmum N15"  = mod_N15_hp,
  #"Hedyosmum CN"   = mod_CN_hp,
  "Hedyosmum Nmass" = mod_N_hp,
  "Hedyosmum Narea"  = mod_Narea_hp,
   "Hedyosmum logNmass"= mod_logNmass_hp,
  "Hedyosmum logNarea"= mod_logNarea_hp,


  "Weinmania iWUE" = mod_iWUE_wl,
  "Weinmania O18"  = mod_O18_wl,
   "Weinmania Nmass" = mod_N_wl,
  "Weinmania Narea"  = mod_Narea_wl,
  "Weinmania logNmass"= mod_logNmass_wl,
  "Weinmania logNarea"= mod_logNarea_wl

)
  #"Weinmania N15"  = mod_N15_wl,
  #"Weinmania CN"   = mod_CN_wl)
# exportar tabla
tab <- modelsummary(
  models,
  output = "data.frame",
  stars = TRUE,
  gof_map = c("nobs", "r.adjusted","adj.squared","r2.marginal","r2.condicional", "aic")
)

# transponer
tab_t <- as.data.frame(t(tab))

# fijar nombres de columna usando la primera fila
colnames(tab_t) <- tab_t[1, ]
tab_t <- tab_t[-1, ]  # quitar fila original de nombres

# mostrar tabla
kable(tab_t)

```
Modelo mejora con la interaccion. Mira AIC.
## MODELOS CON SUB
```{r SUB}

library("sjPlot")
library("sjmisc")
library("sjlabelled")

mod_iWUE_pt <- lmer(`iWUE (μmol/mol)` ~ Sub + (1 | Bloque), data = datos_Pouteria)
mod_O18_pt <- lmer(`δ18O in ‰ vs DeltaSMOW` ~ Sub + (1 | Bloque), data = datos_Pouteria)
mod_N15_pt <- lmer(`δ15N (‰ v.s.AIR)` ~ Sub+ (1 | Bloque), data = datos_Pouteria)
mod_CN_pt <- lmer(`C:N` ~ Sub+ (1 | Bloque), data = datos_Pouteria)
mod_N_pt <- lmer(`[N]` ~ Sub+ (1 | Bloque), data = datos_Pouteria)
mod_Narea_pt <- lmer(`Narea` ~ Sub+ (1 | Bloque), data = datos_Pouteria)
mod_logNarea_pt <- lmer(`logNarea` ~ Sub+ (1 | Bloque), data = datos_Pouteria)
mod_logNmass_pt <- lmer(`logNmass` ~ Sub+ (1 | Bloque), data = datos_Pouteria)

mod_iWUE_cr <- lmer(`iWUE (μmol/mol)` ~ Sub + (1 | Bloque), data = datos_Clarisia)
mod_O18_cr<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ Sub  + (1 | Bloque), data = datos_Clarisia)
mod_N15_cr<- lmer(`δ15N (‰ v.s.AIR)` ~ Sub + (1 | Bloque), data = datos_Clarisia)
mod_CN_cr<- lmer(`C:N` ~Sub+ (1 | Bloque), data = datos_Clarisia)
mod_N_cr <- lmer(`[N]` ~ Sub+ (1 | Bloque), data = datos_Clarisia)
mod_Narea_cr <- lmer(`Narea` ~ Sub+ (1 | Bloque), data = datos_Clarisia)
mod_logNarea_cr <- lmer(`logNarea` ~Sub+ (1 | Bloque), data = datos_Clarisia)
mod_logNmass_cr <- lmer(`logNmass` ~ Sub+ (1 | Bloque), data = datos_Clarisia)

mod_iWUE_al <- lmer(`iWUE (μmol/mol)` ~ Sub + (1 | Bloque), data = datos_al)
mod_O18_al<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ Sub + (1 | Bloque), data = datos_al)
mod_N15_al<- lmer(`δ15N (‰ v.s.AIR)` ~ Sub+  (1 | Bloque), data = datos_al)
mod_CN_al<- lmer(`C:N` ~Sub + (1 | Bloque), data = datos_al)
mod_N_al <- lmer(`[N]` ~ Sub+ (1 | Bloque), data = datos_al)
mod_Narea_al <- lmer(`Narea` ~ Sub+ (1 | Bloque), data = datos_al)
mod_logNarea_al <- lmer(`logNarea` ~ Sub+ (1 | Bloque), data = datos_al)
mod_logNmass_al <- lmer(`logNmass` ~ Sub+ (1 | Bloque), data = datos_al)

mod_iWUE_my <- lmer(`iWUE (μmol/mol)` ~ Sub + (1 | Bloque), data = datos_myrcia)
mod_O18_my<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ Sub+ (1 | Bloque), data = datos_myrcia)
mod_N15_my<- lmer(`δ15N (‰ v.s.AIR)` ~Sub+ (1 | Bloque), data = datos_myrcia)
mod_CN_my<- lmer(`C:N` ~ Sub+ (1 | Bloque), data = datos_myrcia)
mod_N_my <- lmer(`[N]` ~ Sub+ (1 | Bloque), data = datos_myrcia)
mod_Narea_my <- lmer(`Narea` ~ Sub+ (1 | Bloque), data = datos_myrcia)
mod_logNarea_my <- lmer(`logNarea` ~ Sub+ (1 | Bloque), data = datos_myrcia)
mod_logNmass_my <- lmer(`logNmass` ~ Sub+ (1 | Bloque), data = datos_myrcia)

mod_iWUE_hp <- lmer(`iWUE (μmol/mol)` ~Sub+ (1 | Bloque), data = datos_Hp)
mod_O18_hp<- lmer(`δ18O in ‰ vs DeltaSMOW` ~Sub+  (1 | Bloque), data = datos_Hp)
mod_N15_hp<- lmer(`δ15N (‰ v.s.AIR)` ~ Sub+(1 | Bloque), data = datos_Hp)
mod_CN_hp<- lmer(`C:N` ~ Sub + (1 | Bloque), data = datos_Hp)
mod_N_hp <- lmer(`[N]` ~ Sub+ (1 | Bloque), data = datos_Hp)
mod_Narea_hp <- lmer(`Narea` ~ Sub+ (1 | Bloque), data = datos_Hp)
mod_logNarea_hp <- lmer(`logNarea` ~ Sub+ (1 | Bloque), data = datos_Hp)
mod_logNmass_hp <- lmer(`logNmass` ~ Sub+ (1 | Bloque), data = datos_Hp)

mod_iWUE_wl <- lmer(`iWUE (μmol/mol)` ~ Sub + (1 | Bloque), data = datos_Weinmania)
mod_O18_wl<- lmer(`δ18O in ‰ vs DeltaSMOW` ~ Sub + (1 | Bloque), data = datos_Weinmania)
mod_N15_wl<- lmer(`δ15N (‰ v.s.AIR)` ~ Sub + (1 | Bloque), data = datos_Weinmania)
mod_CN_wl<- lmer(`C:N` ~ Sub + (1 | Bloque), data = datos_Weinmania)
mod_N_wl <- lmer(`[N]` ~ Sub+ (1 | Bloque), data = datos_Weinmania)
mod_Narea_wl <- lmer(`Narea` ~ Sub+ (1 | Bloque), data = datos_Weinmania)
mod_logNarea_wl <- lmer(`logNarea` ~ Sub+ (1 | Bloque), data = datos_Weinmania)
mod_logNmass_wl <- lmer(`logNmass` ~ Sub+ (1 | Bloque), data = datos_Weinmania)


## Creates a list with all the models
models  <- list(
  "Pouteria iWUE" = mod_iWUE_pt,
  "Pouteria O18"  = mod_O18_pt,
  #"Pouteria N15"  = mod_N15_pt,
  #"Pouteria CN"   = mod_CN_pt,
  "Pouteria Nmass"=mod_N_pt,
  "Pouteria Narea"=mod_Narea_pt,
  "Pouteria logNmass"= mod_logNmass_pt,
  "Pouteria logNarea"= mod_logNarea_wl,

  "Clarisia iWUE" = mod_iWUE_cr,
  "Clarisia O18"  = mod_O18_cr,
  #"Clarisia N15"  = mod_N15_cr,
  #"Clarisia CN"   = mod_CN_cr,
  "Clarisia Nmass" = mod_N_cr,
  "Clarisia Narea" = mod_Narea_cr,
  "Clar logNmass"= mod_logNmass_cr,
  "Clar logNarea"= mod_logNarea_cr,


  "Alchornea iWUE" = mod_iWUE_al,
  "Alchornea O18"  = mod_O18_al,
  #"Alchornea N15"  = mod_N15_al,
  #"Alchornea CN"   = mod_CN_al,
  "Alchornea Nmass"=mod_N_al,
  "Alchornea Narea"=mod_Narea_al,
    "Alchornea logNmass"= mod_logNmass_al,
  "Alchornea logNarea"= mod_logNarea_al,


  "Myrcia iWUE" = mod_iWUE_my,
  "Myrcia O18"  = mod_O18_my,
  #"Myrcia N15"  = mod_N15_my,
  #"Myrcia CN"   = mod_CN_my,
  "Myrcia Nmass" = mod_N_my,
  "Myrcia Narea"  = mod_Narea_my,
    "Myrcia logNmass"= mod_logNmass_my,
  "Myrcia logNarea"= mod_logNarea_my,


  "Hedyosmum iWUE" = mod_iWUE_hp,
  "Hedyosmum O18"  = mod_O18_hp,
  #"Hedyosmum N15"  = mod_N15_hp,
  #"Hedyosmum CN"   = mod_CN_hp,
  "Hedyosmum Nmass" = mod_N_hp,
  "Hedyosmum Narea"  = mod_Narea_hp,
   "Hedyosmum logNmass"= mod_logNmass_hp,
  "Hedyosmum logNarea"= mod_logNarea_hp,


  "Weinmania iWUE" = mod_iWUE_wl,
  "Weinmania O18"  = mod_O18_wl,
   "Weinmania Nmass" = mod_N_wl,
  "Weinmania Narea"  = mod_Narea_wl,
  "Weinmania logNmass"= mod_logNmass_wl,
  "Weinmania logNarea"= mod_logNarea_wl

)
  #"Weinmania N15"  = mod_N15_wl,
  #"Weinmania CN"   = mod_CN_wl)
# exportar tabla
tab <- modelsummary(
  models,
  output = "data.frame",
  stars = TRUE,
  gof_map = c("nobs", "r.adjusted","adj.squared","r2.marginal","r2.condicional", "aic")
)

# transponer
tab_t <- as.data.frame(t(tab))

# fijar nombres de columna usando la primera fila
colnames(tab_t) <- tab_t[1, ]
tab_t <- tab_t[-1, ]  # quitar fila original de nombres

# mostrar tabla
kable(tab_t)

```
```{r}
library(ggplot2)
library(broom.mixed)
library(dplyr)

coefs <- tidy(
  mod_N_cr,
  effects = "fixed",
  conf.int = TRUE,
  conf.level = 0.95
) %>%
  filter(term != "(Intercept)")

coefs <- coefs %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    )
  )

ggplot(coefs, aes(x = term, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = conf.low, ymax = conf.high),
    width = 0.15
  ) +
  geom_text(
    aes(label = sig, y = conf.high),
    vjust = -0.6,
    size = 5)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  labs(
    x = "Sub",
    y = "Estimate (± 95% CI)"
  ) +
  theme_bw()

coefs$term <- gsub("^Sub", "", coefs$term)

#library(dotwhisker)
#dwplot(mod_iWUE_pt, conf.level = 0.95) +
#  theme_bw()
```


```{r}
plot_model(mod_iWUE_pt, title="Pouteria torta", axis.title="Coefficient", show.values = TRUE, value.offset = .3)
plot_model(mod_iWUE_al, title="Alchornea lojaensis", axis.title="Coefficient", show.values = TRUE, value.offset = .3)
plot_model(mod_iWUE_cr, title="Clarisia racemosa", axis.title="Coefficient", show.values = TRUE, value.offset = .3)
plot_model(mod_iWUE_my, title="Myrcia", axis.title="Coefficient", show.values = TRUE, value.offset = .3)
plot_model(mod_iWUE_hp, title="Hedyosmum", axis.title="Coefficient", show.values = TRUE, value.offset = .3)
plot_model(mod_iWUE_wl, title="Weinmania", axis.title="Coefficient", show.values = TRUE, value.offset = .3)

```




```{r diagnostiscs common }

plot_diagnostics <- function(model, model_name) {
  
  res <- residuals(model)#residual
  fit <- fitted(model) #model
  
  par(mfrow = c(1, 2)) #graph
  
  # Residuals vs Fitted
  plot(fit, res,
       xlab = "Fitted values",
       ylab = "Residuals",
       main = paste(model_name, "\nResiduals vs Fitted"))
  abline(h = 0, col = "red", lty = 2)
  
  # QQ plot
  qqnorm(res, main = paste(model_name, "\nNormal Q-Q"))
  qqline(res, col = "red", lty = 2)
  
  par(mfrow = c(1, 1))
}

#apply the plot_diagnostics_to_all
pdf("model_diagnostics_species.pdf", width = 10, height = 5)

for (nm in names(models)) {
  plot_diagnostics(models[[nm]], nm)
}

dev.off()

```

```{r DHARMA diagnostics}
# Run and plot
for (nm in names(models)) {
  cat("\n", nm, "\n")
  res <- simulateResiduals(models[[nm]], n = 1000)
  plot(res, main = nm)
}

testUniformity(res)
testDispersion(res)
testOutliers(res)

pdf("DHARMa_diagnostics_species.pdf", width = 9, height = 7)

for (nm in names(models)) {
  res <- simulateResiduals(models[[nm]], n = 1000)
  plot(res, main = nm)
}

dev.off()

```







