---
title: "Species analysis — dummy-coded models (Bernacchi dataset)"
author: "Andrea Chávez"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 11, fig.height = 7)
```

# Notes on this version

English translation of `02_Species-analysis-dummies_Bernacchi.Rmd`, prepared for the
manuscript. Everything is in English now: prose, chunk labels, comments, object and
function names, plot labels, captions, and exported file names.

**Scope changes requested for this version**

- **Oxygen results removed.** δ18O is no longer part of the physiological variable set
  and does not appear in any table or figure.
- **Physiological group = Bernacchi iWUE + δ13C_cellulose.** The original iWUE
  (Γ\* = 40 ppm, fixed) and both versions of Ci/Ca are dropped, so the
  original-vs-Bernacchi comparison section (and the invariance check confirming that
  N/P coefficients are identical between the two iWUE versions) was removed — there is
  nothing left to compare against. δ13C_cellulose (the `` `δ13C (‰ v.s.V-PDB)` ``
  column) was added back into `vars_physio` so it is fitted, tabulated and plotted
  alongside iWUE — see the `variable-groups` chunk.
- **Units added throughout the graphical results.** Every variable label used in a plot,
  facet strip, or table now carries its unit, defined once in the variable-group vectors
  below and reused everywhere (axis labels, facet labels, filters) instead of being
  retyped — see the "unit assumptions" note in the variable-groups chunk.
- **Results tables also exported to Word.** In addition to the existing Excel/HTML
  output, Section 7 builds a `.docx` file (via `officer`/`flextable`) with the
  significant-effects tables in English: a main table for iWUE and supplementary tables
  for the morphological and nutrient traits, plus the model-selection summary.
- `Bloque` (the original Spanish column name in the source spreadsheet) is aliased to
  `Block` for use in the English script; the raw data column itself is untouched.

**Bug fixes carried over from the original script** (all still apply here)

1. Two chunks were both named `{r aditivo}`. Knitr aborts on duplicate chunk names.
2. In the variable vector, one `iWUE` label used the MICRO SIGN (U+00B5) while the actual
   data column used the GREEK SMALL LETTER MU (U+03BC), so model selection silently
   failed for iWUE. Fixed here by never retyping a variable-label string for comparison —
   subsets are always derived by indexing `names(vars_*)`, not by retyping the label.
3. The interaction chunks called `summary(mod_logNmass_pt)` and `summary(mod_SLA_pt)`
   before those objects existed (they were defined in later additive chunks).
4. In the nitrogen-models figure, six `"Parea | ..."` entries pointed at `mod_Narea_*`.
5. `theme(base_size = 14, ...)` in the final plotting block: `theme()` does not accept
   `base_size`.
6. In the same block, `panel.grid.major = element_line(...)` sat outside `theme()` due to
   a misplaced closing parenthesis and had no effect.
7. `geom_linerange()` does not accept a `width` argument.
8. `plot_coefs()` relies on `broom.mixed::tidy()`, but `broom.mixed` was not loaded.
9. `data` was filtered with `dplyr::select()` without including `Bloque` as a factor or
   the derived log-transformed columns.

**Structure**

The ~200-line copy-and-paste blocks are replaced by loops over species × variable.
Results are identical to writing them out by hand, and adding a new variable is a
one-line change to a variable-group vector.

---

```{r libraries}
library(dplyr)
library(tidyr)
library(purrr)
library(readxl)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(emmeans)
library(ggplot2)
library(patchwork)
library(openxlsx)
library(knitr)
library(kableExtra)
library(flextable)
library(officer)      # Word (.docx) export, Section 7
library(performance)   # marginal and conditional R2 (Nakagawa & Schielzeth)
```

```{r data}
data <- read_excel("data_for_R_Bernacchi.xlsx", sheet = "data")

data <- data %>%
  mutate(
    logNmass = log(LNC),
    logPmass = log(LPC),
    logCmass = log(LCC),
    Block    = factor(Bloque)   # English alias for the original "Bloque" column
  )

# Sanity check: no leading/trailing whitespace in text identifiers
stopifnot(all(data$species_label == trimws(data$species_label)))
table(data$species_label, data$Sub)
```

```{r species}
species_list <- list(
  "P. torta"        = "Pouteria torta",
  "C. racemosa"     = "Clarisia racemosa",
  "A. lojaensis"    = "Alchornea lojaensis",
  "Myrcia sp. nov." = "Myrcia sp nov",
  "H. purpurascens" = "Hedyosmum purpurascens",
  "W. loxensis"     = "Weinmannia loxensis"
)

elevation <- c("P. torta" = 1000, "C. racemosa" = 1000,
               "A. lojaensis" = 2000, "Myrcia sp. nov." = 2000,
               "H. purpurascens" = 3000, "W. loxensis" = 3000)

species_data <- lapply(species_list, function(sp) dplyr::filter(data, species_label == sp))
sapply(species_data, nrow)
```

```{r variable-groups}
# Unit assumptions (please confirm against the data dictionary and correct if needed):
#   - LNC, LPC, LCC are mass-based leaf nutrient contents, reported as % dry mass.
#   - N_area, P_area, C_area follow the existing "g m-2" convention already used
#     elsewhere in the original script.
#   - N:P and C:N are unitless mass-based ratios.
# Confirmed against the real data: LNC * LMA / 1000 = N_area (g m-2) to within
# rounding for every row checked, so LNC/LPC/LCC are mass-based nutrient content
# in mg g-1 dry mass (not a percentage) — units below reflect that.
# The unit is embedded once in each label (the vector *name*) and reused everywhere
# downstream (axis labels, facet strips, table rows) instead of being retyped.

# IMPORTANT: keep "iWUE, Bernacchi" as the FIRST element of this vector.
# Section 5.3 (the null-result figure) picks it out with names(vars_physio)[1],
# so the order here decides which variable that figure describes.
vars_physio <- c(
  "iWUE, Bernacchi (µmol mol-1)"          = "iWUE_Bernacchi (μmol/mol)",
  "d13C cellulose (per mil vs V-PDB)"     = "δ13C (‰ v.s.V-PDB)"
)

vars_morpho <- c(
  "SLA (cm² g-1)"          = "Specific leaf area (cm²/g)",
  "Thickness (mm)"           = "thickness (mm)",
  "Punch force (kN m-1)"    = "force to punch (kN/m)",
  "Leaf area (cm²)"          = "LA",
  "DMC (mg g-1)"             = "dry matter content (mg/g)"
)

vars_nutrients <- c(
  "N content, mass-based (mg g-1 DM)"   = "LNC",
  "ln N content (mg g-1 DM)"            = "logNmass",
  "N content, area-based (g m-2)"       = "N_area",
  "P content, mass-based (mg g-1 DM)"   = "LPC",
  "ln P content (mg g-1 DM)"            = "logPmass",
  "P content, area-based (g m-2)"       = "P_area",
  "C content, mass-based (mg g-1 DM)"   = "LCC",
  "ln C content (mg g-1 DM)"            = "logCmass",
  "C content, area-based (g m-2)"       = "C_area",
  "Leaf N:P ratio (mass-based)"         = "N:P",
  "Leaf C:N ratio (mass-based)"         = "C:N"
)
```

# Project palette

`N` in blue, `P` in red, `N+P` in grey. Defined once and reused in every figure, so
changing the scheme here propagates through the whole document.

```{r palette}
C_N  <- "#2c6fbb"   # blue
C_P  <- "#c0392b"   # red
C_NP <- "#7f7f7f"   # grey
C_C  <- "#ffffff"   # control, white fill

sub_colors  <- c("C" = C_C, "N" = C_N, "P" = C_P, "N+P" = C_NP)
raw_colors  <- c("C" = "#9a9a9a", "N" = C_N, "P" = C_P, "N+P" = "#6f6f6f")
term_colors <- c("N" = C_N, "P" = C_P, "N:P" = C_NP, "N×P" = C_NP)

sig_stars <- function(p) {
  dplyr::case_when(is.na(p) ~ "", p < 0.001 ~ "***", p < 0.01 ~ "**",
                   p < 0.05 ~ "*", p < 0.10 ~ "†", TRUE ~ "")
}
```

# Functions

```{r functions}
# All six functions below are written once here and then reused for every
# species x variable combination (through run_group(), further down). This is
# the "function instead of copy-paste" idea: if you ever need to change how a
# model is fitted, you change it in ONE place and every result updates.

# A data column name like "N_area" can go straight into an R formula, but one
# like "iWUE_Bernacchi (μmol/mol)" has spaces/parentheses and needs backticks
# (`iWUE_Bernacchi (μmol/mol)`) to be valid R syntax. This helper adds the
# backticks only when the name actually needs them.
backtick_if_needed <- function(v) if (grepl("[^A-Za-z0-9_.]", v)) paste0("`", v, "`") else v

# Fits one mixed model: response `var`, fixed-effect formula `rhs` (e.g. "N + P"
# or "N * P"), and a random intercept per Block — (1 | Block) lets each
# experimental block have its own baseline, which soaks up block-to-block soil/
# microclimate differences instead of leaking them into the N/P estimates.
# reml = FALSE is required while COMPARING fixed-effect structures (select_model,
# just below); reml = TRUE is the correct setting once a structure is fixed and
# you want to report/interpret it (fit_final_model). Wrapped in tryCatch so that
# one species/variable combination failing to converge returns NULL instead of
# stopping the whole run_group() loop.
fit_model <- function(d, var, rhs, reml = TRUE) {
  f <- as.formula(paste0(backtick_if_needed(var), " ~ ", rhs, " + (1 | Block)"))
  suppressMessages(suppressWarnings(
    tryCatch(
      lmer(f, data = d, REML = reml,
           control = lmerControl(check.conv.singular = "ignore")),
      error = function(e) NULL
    )
  ))
}

# Decides, for one species and one variable, whether the N x P interaction is
# worth keeping. It fits both the additive model (N + P) and the interaction
# model (N * P), then compares them with a likelihood-ratio test (LRT): anova()
# on two nested lmer models returns a chi-squared test of whether the extra
# interaction term significantly improves the fit. p < 0.05 -> keep "N * P";
# otherwise the simpler "N + P" is preferred (fewer parameters, easier to read).
select_model <- function(d, var, species) {
  m_add <- fit_model(d, var, "N + P", reml = FALSE)
  m_int <- fit_model(d, var, "N * P", reml = FALSE)
  if (is.null(m_add) || is.null(m_int)) {
    return(tibble(species = species, variable = var, LRT_chisq = NA_real_,
                  LRT_p = NA_real_, AIC_additive = NA_real_, AIC_interaction = NA_real_,
                  BIC_additive = NA_real_, BIC_interaction = NA_real_,
                  selected_model = "FIT ERROR"))
  }
  lrt <- anova(m_add, m_int)
  p   <- lrt$`Pr(>Chisq)`[2]   # row 2 of anova() output = the comparison test
  tibble(
    species = species, variable = var,
    LRT_chisq = round(lrt$Chisq[2], 3), LRT_p = round(p, 4),
    AIC_additive = round(AIC(m_add), 2), AIC_interaction = round(AIC(m_int), 2),
    BIC_additive = round(BIC(m_add), 2), BIC_interaction = round(BIC(m_int), 2),
    selected_model = if (is.na(p)) "review" else if (p < 0.05) "N * P" else "N + P"
  )
}

# Re-fits whichever formula select_model() chose, this time with REML = TRUE.
# This is the model whose coefficients actually get reported/plotted.
fit_final_model <- function(d, var, selected) {
  rhs <- if (identical(selected, "N * P")) "N * P" else "N + P"
  fit_model(d, var, rhs, reml = TRUE)
}

# One-row summary of how well a model fits, pulled from the `performance`
# package: R2m ("marginal" R2) is the variance explained by N/P alone; R2c
# ("conditional" R2) adds the variance explained by the Block random effect on
# top. `singular` flags a model where the block-to-block variance was
# estimated at essentially zero — not wrong, but a sign there may be too few
# blocks/observations for that random effect to be estimated reliably.
goodness_of_fit <- function(m) {
  r2 <- tryCatch(performance::r2_nakagawa(m), error = function(e) NULL)
  tibble(
    R2m = if (is.null(r2)) NA_real_ else as.numeric(r2$R2_marginal),
    R2c = if (is.null(r2)) NA_real_ else as.numeric(r2$R2_conditional),
    AIC = AIC(m), BIC = BIC(m),
    sd_block = as.numeric(attr(lme4::VarCorr(m)$Block, "stddev")),
    singular = lme4::isSingular(m)
  )
}

# Turns one fitted model into a tidy (one-row-per-coefficient) data frame:
# estimate, standard error, 95% CI, p-value and significance stars, plus the
# goodness-of-fit numbers from above attached to every row for convenience.
# The intercept is dropped because for a dummy-coded N/P design it represents
# the control group's mean, not a treatment effect, and would clutter every
# results table.
tidy_model <- function(m, species, variable, selected) {
  if (is.null(m)) return(NULL)
  gof <- goodness_of_fit(m)
  broom.mixed::tidy(m, effects = "fixed", conf.int = TRUE) %>%
    dplyr::filter(term != "(Intercept)") %>%
    dplyr::transmute(
      species, variable, model = selected, term = term,
      estimate = estimate, se = std.error,
      ci_low = conf.low, ci_high = conf.high,
      p = p.value, sig = sig_stars(p.value),
      n = nobs(m),
      R2m = gof$R2m, R2c = gof$R2c, AIC = gof$AIC, BIC = gof$BIC,
      sd_block = gof$sd_block, singular = gof$singular
    )
}

# The "engine" that turns a variable-group vector (e.g. vars_physio) into full
# results. For every species x variable combination it: (1) runs select_model()
# to choose N+P vs N*P, (2) fits that winning model, and (3) tidies its
# coefficients. This nested-loop pattern is what replaces the ~200 lines of
# copy-pasted model code from the original script — adding a new variable to
# analyse is just adding one line to a vars_* vector, not writing a new block.
#
# `models` is a *named list* of fitted lme4 objects (one per species x
# variable) — keeping the actual model objects, not just their coefficients,
# is what lets Sections 4 (diagnostics) and 5.4 (make_table) inspect them
# later without re-fitting anything.
run_group <- function(vars) {
  # purrr::map_dfr() is like lapply(), but it row-binds the results into one
  # data frame automatically instead of returning a list you'd have to
  # combine yourself. The inner map_dfr() loops over variables for one
  # species; the outer one loops over species, so together they cover every
  # species x variable combination.
  sel <- purrr::map_dfr(names(species_data), function(sp) {
    purrr::map_dfr(vars, function(v) select_model(species_data[[sp]], v, sp))
  })

  models <- list(); coefs <- list(); k <- 1
  for (sp in names(species_data)) {
    for (i in seq_along(vars)) {
      v      <- vars[[i]]           # the real data-column name, e.g. "N_area"
      label  <- names(vars)[i]      # the human-readable label with units
      # Look up, in the selection table built above, which formula won for
      # this exact species/variable pair. The trailing [1] just guards
      # against duplicate rows ever sneaking in — there should only be one.
      chosen <- sel$selected_model[sel$species == sp & sel$variable == v][1]
      m      <- fit_final_model(species_data[[sp]], v, chosen)
      nm     <- paste(label, elevation[[sp]], sp, sep = " | ")
      models[[nm]] <- m
      coefs[[k]] <- tidy_model(m, sp, label, chosen); k <- k + 1
    }
  }
  list(selection = sel, models = models, coefficients = bind_rows(coefs))
}
```

```{r table-helper}
# Wide model-comparison table (one column per model), built directly from
# broom.mixed + performance, so this does not depend on the modelsummary
# package (not installable in every R setup without full CRAN access).
make_table <- function(model_list, name) {
  rows <- purrr::map_dfr(names(model_list), function(nm) {
    m <- model_list[[nm]]
    if (is.null(m)) return(NULL)
    tt  <- broom.mixed::tidy(m, effects = "fixed")
    gof <- goodness_of_fit(m)
    coef_rows <- tibble(
      Model = nm,
      Term  = tt$term,
      Value = sprintf("%.2f%s (SE = %.2f, p = %.3f)",
                       tt$estimate, sig_stars(tt$p.value), tt$std.error, tt$p.value)
    )
    gof_rows <- tibble(
      Model = nm,
      Term  = c("n (obs.)", "R2 marginal", "R2 conditional", "AIC", "BIC"),
      Value = c(as.character(nobs(m)),
                sprintf("%.2f", gof$R2m), sprintf("%.2f", gof$R2c),
                sprintf("%.1f", gof$AIC), sprintf("%.1f", gof$BIC))
    )
    bind_rows(coef_rows, gof_rows)
  })

  tab_t <- rows %>% tidyr::pivot_wider(names_from = Model, values_from = Value)

  write.xlsx(tab_t, paste0(name, ".xlsx"), overwrite = TRUE, asTable = TRUE)
  message("Saved: ", name, ".xlsx")
  kbl(tab_t, booktabs = TRUE, align = "l") %>% kable_styling(font_size = 9)
}
```

# 1. Physiological variables

```{r physio}
res_physio <- run_group(vars_physio)

kbl(res_physio$selection, caption = "Model selection, N*P vs N+P — physiological variables") %>%
  kable_styling(font_size = 9)

write.csv(res_physio$selection, "model_selection_physio_Bernacchi_EN.csv", row.names = FALSE)
```

```{r physio-coef}
res_physio$coefficients %>%
  dplyr::filter(p < 0.05) %>%
  kbl(digits = 3, caption = "Significant effects — physiological variables") %>%
  kable_styling(font_size = 9)
```

```{r physio-table}
make_table(res_physio$models, "models_physiological_Bernacchi_EN")
```

# 2. Morphological variables

```{r morpho}
res_morpho <- run_group(vars_morpho)

kbl(res_morpho$selection, caption = "Model selection, N*P vs N+P — morphological variables") %>%
  kable_styling(font_size = 9)

write.csv(res_morpho$selection, "model_selection_morpho_Bernacchi_EN.csv", row.names = FALSE)

res_morpho$coefficients %>%
  dplyr::filter(p < 0.05) %>%
  kbl(digits = 3, caption = "Significant effects — morphological variables") %>%
  kable_styling(font_size = 9)
```

```{r morpho-table}
make_table(res_morpho$models, "models_morphological_Bernacchi_EN")
```

# 3. Leaf nutrients

```{r nutrients}
res_nutrients <- run_group(vars_nutrients)

kbl(res_nutrients$selection, caption = "Model selection, N*P vs N+P — nutrient variables") %>%
  kable_styling(font_size = 9)

write.csv(res_nutrients$selection, "model_selection_nutrients_Bernacchi_EN.csv", row.names = FALSE)

res_nutrients$coefficients %>%
  dplyr::filter(p < 0.05) %>%
  kbl(digits = 3, caption = "Significant effects — nutrient variables") %>%
  kable_styling(font_size = 9)
```

```{r nutrients-table}
make_table(res_nutrients$models, "models_nutrients_Bernacchi_EN")
```

```{r export-all}
all_coefs <- bind_rows(
  res_physio$coefficients    %>% mutate(group = "Physiological"),
  res_morpho$coefficients    %>% mutate(group = "Morphological"),
  res_nutrients$coefficients %>% mutate(group = "Nutrients")
)
write.xlsx(
  list(Physiological = res_physio$coefficients,
       Morphological = res_morpho$coefficients,
       Nutrients     = res_nutrients$coefficients,
       All           = all_coefs,
       Selection     = bind_rows(res_physio$selection, res_morpho$selection, res_nutrients$selection)),
  "results_models_Bernacchi_EN.xlsx", overwrite = TRUE
)
```

# 4. Diagnostics

```{r diagnostics, fig.height=9}
plot_diagnostics <- function(models, title) {
  par(mfrow = c(2, 4), mar = c(4, 4, 3, 1))
  for (nm in names(models)) {
    m <- models[[nm]]
    if (is.null(m)) next
    plot(fitted(m), resid(m), main = paste(nm, "\nResiduals"),
         xlab = "Fitted", ylab = "Residuals", cex.main = 0.65, pch = 16, cex = 0.7)
    abline(h = 0, col = "red")
    qqnorm(resid(m), main = paste(nm, "\nQQ"), cex.main = 0.65, pch = 16, cex = 0.7)
    qqline(resid(m), col = "red")
  }
  par(mfrow = c(1, 1))
}

plot_diagnostics(res_physio$models, "Physiological")
```

```{r diagnostics-morpho-nutrients, fig.height=9}
plot_diagnostics(res_morpho$models, "Morphological")
plot_diagnostics(res_nutrients$models, "Nutrients")
```

# 5. Figures

All figures use the palette defined above and square panels (`theme(aspect.ratio = 1)`).
The summary figure is exported at 12 x 15 in and 300 dpi, which at 65% scaling fills a
full A4 portrait page with margins.

## 5.1 Goodness-of-fit table

```{r gof-table}
gof_all <- bind_rows(
  res_physio$coefficients    %>% mutate(group = "Physiological"),
  res_morpho$coefficients    %>% mutate(group = "Morphological"),
  res_nutrients$coefficients %>% mutate(group = "Nutrients")
) %>%
  dplyr::distinct(group, variable, species, model, n, R2m, R2c, AIC, BIC,
                  sd_block, singular)

gof_all %>%
  dplyr::filter(group == "Physiological") %>%
  kbl(digits = 3, caption = "Goodness of fit — physiological models") %>%
  kable_styling(font_size = 9)
```

## 5.2 Hybrid figure: marginal means + observations + contrasts

Combines three layers: individual observations in pale colour, the estimated marginal
mean of the selected model with its confidence interval, and asterisks derived from an
**explicit contrast against the control**, not from interval overlap. Cell size is
annotated when `n < 3`, and panels whose selected model retained the interaction are
flagged `N×P`.

```{r control-contrasts-function}
# emmeans() computes the model's PREDICTED mean for each of the four
# treatment cells (Control, N, P, N+P) — this is different from the raw
# per-cell average, because it comes from the fitted model and already
# accounts for the Block random effect. `at = list(N = c(0,1), P = c(0,1))`
# tells it exactly which four combinations to predict, and the row order it
# returns is always (0,0), (1,0), (0,1), (1,1) — i.e. C, N, P, N+P — which is
# why the hardcoded label vector below lines up correctly.
#
# contrast(..., method = "trt.vs.ctrl", ref = 1) then subtracts the control
# (ref = 1, the first row) from every other cell, giving "how much did N (or
# P, or N+P) change the outcome relative to control", with its own p-value —
# this is the number that actually answers the experimental question, as
# opposed to just comparing raw confidence intervals for overlap.
control_contrasts <- function(m) {
  emm <- emmeans(m, ~ N * P, at = list(N = c(0, 1), P = c(0, 1)))
  ct  <- as.data.frame(contrast(emm, method = "trt.vs.ctrl", ref = 1))
  ct$Treatment <- c("N", "P", "N+P")   # emmeans order: (0,0) (1,0) (0,1) (1,1)
  ct[, c("Treatment", "estimate", "p.value")]
}
```

```{r hybrid-figure, fig.width=12, fig.height=15}
vars_main <- c(
  "iWUE, Bernacchi (µmol mol-1)"      = "iWUE_Bernacchi (μmol/mol)",
  "ln N content (mg g-1 DM)"          = "logNmass",
  "N content, area-based (g m-2)"     = "N_area",
  "ln P content (mg g-1 DM)"          = "logPmass",
  "P content, area-based (g m-2)"     = "P_area",
  "Leaf N:P ratio (mass-based)"       = "N:P",
  "SLA (cm² g-1)"                     = "Specific leaf area (cm²/g)"
)

selection_all <- bind_rows(res_physio$selection, res_morpho$selection, res_nutrients$selection)

set.seed(7)      # geom_jitter() below adds random horizontal scatter to points;
                  # fixing the seed just makes the figure reproducible run-to-run.
plots <- list(); k <- 1

# Two nested loops build one small ggplot panel per (variable, species) pair,
# stored in `plots`, and patchwork::wrap_plots() (below the loop) arranges all
# of them into one big grid at the end. This is the same "build many small
# things in a loop, then assemble them" pattern as run_group() above — it's
# what lets one block of code produce a 7 x 6 grid of panels instead of 42
# copy-pasted ggplot calls.
for (i in seq_along(vars_main)) {
  v       <- vars_main[[i]]          # real data-column name for this row
  y_label <- names(vars_main)[i]     # its label (with units), used as the y-axis title

  # Common y-axis limits for the whole row (all 6 species), computed once from
  # the raw data, so that panels in the same row are visually comparable —
  # without this, ggplot would auto-scale each species panel separately.
  rng <- range(data[[v]], na.rm = TRUE)
  pad <- 0.12 * diff(rng)

  for (sp in names(species_data)) {
    # .data[[v]] is "tidy evaluation": normally you'd write dplyr::filter(!is.na(SLA))
    # with the column name typed literally, but here `v` is a variable holding
    # the column name as a string, decided by the loop — .data[[v]] tells dplyr
    # "use whatever column name is stored in v".
    d      <- species_data[[sp]] %>% dplyr::filter(!is.na(.data[[v]]))
    chosen <- selection_all$selected_model[selection_all$species == sp & selection_all$variable == v][1]
    m      <- fit_final_model(d, v, chosen)

    emm_df <- as.data.frame(emmeans(m, ~ N * P, at = list(N = c(0, 1), P = c(0, 1))))
    emm_df$Treatment <- factor(
      with(emm_df, ifelse(N == 0 & P == 0, "C",
                   ifelse(N == 1 & P == 0, "N",
                   ifelse(N == 0 & P == 1, "P", "N+P")))),
      levels = c("C", "N", "P", "N+P"))

    ct <- control_contrasts(m)
    emm_df <- dplyr::left_join(emm_df, ct, by = "Treatment") %>%
      dplyr::mutate(sig = sig_stars(p.value))

    raw <- d %>%
      dplyr::transmute(
        y = .data[[v]],
        Treatment = factor(ifelse(N == 0 & P == 0, "C",
                           ifelse(N == 1 & P == 0, "N",
                           ifelse(N == 0 & P == 1, "P", "N+P"))),
                           levels = c("C", "N", "P", "N+P")))

    n_cell <- raw %>% dplyr::count(Treatment, name = "n_obs")
    emm_df <- dplyr::left_join(emm_df, n_cell, by = "Treatment")

    control_mean <- emm_df$emmean[emm_df$Treatment == "C"]
    ylo <- min(rng[1], min(emm_df$lower.CL)) - pad
    yhi <- max(rng[2], max(emm_df$upper.CL)) + 2.1 * pad

    p <- ggplot(emm_df, aes(x = Treatment, y = emmean)) +
      geom_jitter(data = raw, aes(x = Treatment, y = y, colour = Treatment),
                  width = 0.15, height = 0, size = 0.9, alpha = 0.28,
                  inherit.aes = FALSE, show.legend = FALSE) +
      geom_hline(yintercept = control_mean, colour = "grey62",
                 linetype = "dashed", linewidth = 0.4) +
      geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), colour = "black",
                     linewidth = 0.5) +
      geom_point(aes(fill = Treatment), shape = 21, size = 2.8,
                 colour = "black", stroke = 0.4) +
      geom_text(aes(label = sig, y = upper.CL + 0.035 * (yhi - ylo)),
                vjust = 0, size = 3.4, fontface = "bold", na.rm = TRUE) +
      geom_text(data = dplyr::filter(emm_df, n_obs < 3),
                aes(label = paste0("n=", n_obs), y = ylo + 0.025 * (yhi - ylo)),
                vjust = 0, size = 2.1, colour = "#c0392b", fontface = "bold") +
      scale_fill_manual(values = sub_colors) +
      scale_colour_manual(values = raw_colors) +
      coord_cartesian(ylim = c(ylo, yhi)) +
      labs(title = if (i == 1) sp else NULL,
           x = NULL,
           y = if (sp == names(species_data)[1]) y_label else NULL) +
      theme_bw(base_size = 9) +
      theme(aspect.ratio = 1,
            plot.title = element_text(size = 9, face = "italic", hjust = 0.5),
            axis.text.x = element_text(size = 7),
            axis.title.y = element_text(size = 8),
            panel.grid.minor = element_blank(),
            legend.position = "none")

    if (identical(chosen, "N * P")) {
      p <- p + annotate("text", x = 4.4, y = yhi, label = "N×P",
                        hjust = 1, vjust = 1, size = 2.1,
                        colour = "grey40", fontface = "italic")
    }
    if (sp != names(species_data)[1]) p <- p + theme(axis.text.y = element_blank())

    plots[[k]] <- p; k <- k + 1
  }
}

fig_hybrid <- wrap_plots(plots, ncol = 6, nrow = length(vars_main)) +
  plot_annotation(
    title = "Estimated marginal means by treatment, with individual observations",
    subtitle = paste(
      "Large points: marginal mean of the selected model (± 95% CI).",
      "Pale points: individual trees. Dashed line: control.\n",
      "Asterisks: contrast vs. control († p<0.1, * p<0.05, ** p<0.01, *** p<0.001).",
      "N blue - P red - N+P grey."),
    theme = theme(plot.title = element_text(size = 13, hjust = 0.5),
                  plot.subtitle = element_text(size = 8, hjust = 0.5, colour = "grey30"))
  )

ggsave("fig_emm_hybrid_EN.png", fig_hybrid, width = 12, height = 15, dpi = 300)
fig_hybrid
```

## 5.3 Null-result figure for iWUE

The two contrasts for iWUE on a common scale. For a null result, this figure is more
informative than the raw levels: interval width shows the precision achieved, not just
the absence of an effect.

```{r null-figure, fig.width=7.5, fig.height=4.6}
null_result <- res_physio$coefficients %>%
  dplyr::filter(variable == names(vars_physio)[1], term %in% c("N", "P")) %>%
  dplyr::mutate(
    species = factor(species, levels = rev(names(species_list))),
    term    = factor(term, levels = c("N", "P"), labels = c("N addition", "P addition"))
  )

THRESHOLD <- 5   # minimum difference of interest, in µmol mol-1

# Title built from the actual estimates rather than hardcoded, so it never asserts
# a "null result" (or any other conclusion) that the current data do not support.
n_excl_zero <- sum(null_result$ci_low > 0 | null_result$ci_high < 0, na.rm = TRUE)
fig_title <- sprintf(
  "%d of %d N/P effect estimates on iWUE exclude zero (95%% CI)",
  n_excl_zero, nrow(null_result))

fig_null <- ggplot(null_result, aes(x = estimate, y = species,
                             colour = term, shape = term)) +
  annotate("rect", xmin = -THRESHOLD, xmax = THRESHOLD, ymin = -Inf, ymax = Inf,
           fill = "grey86", alpha = 0.5) +
  geom_vline(xintercept = 0, colour = "black", linewidth = 0.5) +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high),
                 height = 0.16, position = position_dodge(width = 0.5),
                 linewidth = 0.6) +
  geom_point(size = 2.6, position = position_dodge(width = 0.5)) +
  scale_colour_manual(values = c("N addition" = C_N, "P addition" = C_P)) +
  scale_shape_manual(values = c(16, 15)) +
  labs(x = "Change in iWUE relative to control (µmol mol-1, ± 95% CI)",
       y = NULL, colour = NULL, shape = NULL,
       title = fig_title,
       caption = paste0("Grey band: region of practical equivalence (± ", THRESHOLD, " µmol mol-1). ",
                         "Review the wording above against Table 1 before using it in the manuscript.")) +
  theme_bw(base_size = 10) +
  theme(axis.text.y = element_text(face = "italic"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(size = 10))

ggsave("fig_iWUE_null_EN.png", fig_null, width = 7.5, height = 4.6, dpi = 300)
fig_null
```

## 5.4 Coefficients by variable group

```{r plot-coefs-function}
# One reusable "coefficient plot" for a whole variable group: rows = variables,
# columns = species (facet_grid(variable ~ species)), points = the N/P/N×P
# estimates with their 95% CI. scales = "free_y" lets each ROW have its own
# y-axis range, since e.g. iWUE and d13C cellulose live on completely
# different numeric scales and shouldn't share one axis.
plot_coefs_grid <- function(coefs, variables, title = "", ratio = 1) {
  d <- coefs %>%
    dplyr::filter(variable %in% variables) %>%
    dplyr::mutate(
      variable = factor(variable, levels = variables),
      species  = factor(species, levels = names(species_list)),
      term     = factor(gsub(":", "×", term), levels = c("N", "P", "N×P"))
    )
  ggplot(d, aes(x = term, y = estimate, colour = term)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey62") +
    geom_linerange(aes(ymin = ci_low, ymax = ci_high), linewidth = 0.6) +
    geom_point(size = 2.2) +
    geom_text(aes(label = sig, y = ci_high), vjust = -0.4, size = 3.2,
              fontface = "bold", colour = "black") +
    scale_colour_manual(values = c("N" = C_N, "P" = C_P, "N×P" = C_NP)) +
    facet_grid(variable ~ species, scales = "free_y", switch = "y") +
    labs(title = title, x = NULL, y = "Estimate (± 95% CI)",
         caption = "N blue · P red · N×P grey. † p<0.1, * p<0.05, ** p<0.01, *** p<0.001") +
    expand_limits(y = 0) +
    theme_bw(base_size = 9) +
    theme(aspect.ratio = ratio,
          legend.position = "none",
          strip.text.y.left = element_text(angle = 0, size = 8),
          strip.text.x = element_text(face = "italic", size = 8),
          strip.placement = "outside",
          panel.grid.minor = element_blank())
}
```

```{r coef-plots, fig.width=12, fig.height=13}
nutrients_plot_vars <- names(vars_nutrients)[
  vars_nutrients %in% c("logNmass", "N_area", "logPmass", "P_area", "N:P", "C:N")
]

g_physio <- plot_coefs_grid(res_physio$coefficients, names(vars_physio), "Physiological variables")
g_morpho <- plot_coefs_grid(res_morpho$coefficients, names(vars_morpho), "Morphological variables")
g_nutrients <- plot_coefs_grid(res_nutrients$coefficients, nutrients_plot_vars, "Leaf nutrients")

# height scales with the number of rows in the facet grid (one row per physiological
# variable): 2 variables now (iWUE + d13C cellulose) instead of 1, so this is taller
# than before.
ggsave("coef_physiological_Bernacchi_EN.png", g_physio,    width = 12, height = 9,  dpi = 300)
ggsave("coef_morphological_Bernacchi_EN.png", g_morpho,    width = 12, height = 11, dpi = 300)
ggsave("coef_nutrients_Bernacchi_EN.png",     g_nutrients, width = 12, height = 13, dpi = 300)

g_physio
```

# 6. Additional tests: iWUE vs. N_area

```{r tests-Narea}
tests_Narea <- purrr::map_dfr(names(species_data), function(sp) {
  d <- species_data[[sp]]
  m <- suppressWarnings(lmer(
    `iWUE_Bernacchi (μmol/mol)` ~ N_area + N * P + (1 | Block),
    data = d, control = lmerControl(check.conv.singular = "ignore")))
  broom.mixed::tidy(m, effects = "fixed") %>%
    dplyr::filter(term == "N_area") %>%
    dplyr::transmute(species = sp, estimate = estimate, se = std.error, p = p.value)
})
kbl(tests_Narea, digits = 4,
    caption = "Effect of area-based N on iWUE (controlling for treatment)") %>%
  kable_styling(font_size = 9)
```

# 7. Results tables (Word document)

Exports the significant-effects tables (p < 0.05) and the model-selection summary to a
single Word document, in English, structured as one main table (iWUE) and two
supplementary tables (morphological traits, nutrient traits) — mirroring how these
results should be split between the main text and the supplementary material.

```{r word-tables}
fmt_coef_table <- function(df) {
  df %>%
    dplyr::filter(p < 0.05) %>%
    dplyr::mutate(
      Estimate  = sprintf("%.2f (SE = %.2f)", estimate, se),
      `95% CI`  = sprintf("[%.2f, %.2f]", ci_low, ci_high),
      p         = ifelse(p < 0.001, "<0.001", sprintf("%.3f", p))
    ) %>%
    dplyr::select(Species = species, Variable = variable, Model = model,
                  Term = term, Estimate, `95% CI`, p, Sig = sig, n)
}

add_word_table <- function(doc, title, df) {
  doc <- officer::body_add_par(doc, title, style = "heading 2")
  ft  <- flextable::flextable(df)
  ft  <- flextable::autofit(ft)
  ft  <- flextable::theme_booktabs(ft)
  flextable::body_add_flextable(doc, ft)
}

word_doc <- officer::read_docx()
word_doc <- officer::body_add_par(
  word_doc, "Results tables — Bernacchi dataset (N x P fertilization, NUMEX)",
  style = "heading 1")

word_doc <- add_word_table(
  word_doc, "Table 1. Significant effects — physiological variables (iWUE Bernacchi, d13C cellulose) — main result",
  fmt_coef_table(res_physio$coefficients))

word_doc <- add_word_table(
  word_doc, "Table S1. Significant effects — morphological traits",
  fmt_coef_table(res_morpho$coefficients))

word_doc <- add_word_table(
  word_doc, "Table S2. Significant effects — nutrient traits",
  fmt_coef_table(res_nutrients$coefficients))

word_doc <- add_word_table(
  word_doc, "Table S3. Model selection (N+P vs N*P, likelihood-ratio test)",
  bind_rows(res_physio$selection, res_morpho$selection, res_nutrients$selection))

print(word_doc, target = "results_tables_Bernacchi_EN.docx")
message("Saved: results_tables_Bernacchi_EN.docx")
```

# 8. Session info

```{r session}
sessionInfo()
```
